---
title: "Park Sizes in Vancouver"
author: "Qibing He"
date: 2026-09-25
---

This post examines the distribution of public park sizes in Vancouver. I will use R to inspect park areas, calculate summary statistics, identify the largest parks, and visualize the distribution of park sizes.

Data: [Parks](https://opendata.vancouver.ca/explore/dataset/parks/), City of Vancouver Open Data Portal. The dataset is provided under the Open Government Licence – Vancouver.

## Setup

The following chunk loads the R package used for visualization.

```{r}
#| label: setup-r
#| message: false
#| warning: false

library(ggplot2)
```

## Load and inspect the data

First, I load the parks dataset and inspect its dimensions and selected columns.

```{r}
#| label: load-parks-r

parks <- read.csv(
  "../../data/parks.csv",
  sep = ";",
  stringsAsFactors = FALSE
)

cat(
  "The dataset contains",
  nrow(parks),
  "parks and",
  ncol(parks),
  "columns.\n"
)

head(parks[c("Name", "NeighbourhoodName", "Hectare")])
```

## Check and clean the data

I check the variables used in the analysis for missing values before calculating park-size statistics.

```{r}
#| label: check-missing-r

analysis_columns <- c(
  "Name",
  "NeighbourhoodName",
  "Hectare"
)

colSums(is.na(parks[analysis_columns]))
```

```{r}
#| label: prepare-parks-r

parks_clean <- parks[
  complete.cases(parks[analysis_columns]),
]

parks_clean$Hectare <- as.numeric(
  parks_clean$Hectare
)

parks_clean <- parks_clean[
  !is.na(parks_clean$Hectare),
]

cat(
  nrow(parks_clean),
  "parks remain after cleaning.\n"
)
```

## Summary of park sizes

I calculate several summary statistics to understand the typical park size and the overall range of park areas.

```{r}
#| label: summarize-park-sizes

size_summary <- data.frame(
  Statistic = c(
    "Mean",
    "Median",
    "Minimum",
    "Maximum"
  ),
  Hectares = round(
    c(
      mean(parks_clean$Hectare),
      median(parks_clean$Hectare),
      min(parks_clean$Hectare),
      max(parks_clean$Hectare)
    ),
    2
  )
)

size_summary
```

## Largest parks

The following table identifies the ten largest parks in the dataset.

```{r}
#| label: largest-parks

largest_parks <- parks_clean[
  order(
    parks_clean$Hectare,
    decreasing = TRUE
  ),
  c(
    "Name",
    "NeighbourhoodName",
    "Hectare"
  )
]

head(largest_parks, 10)
```

## Visualize the largest parks

The following chart compares the areas of the ten largest parks in the dataset.

```{r}
#| label: fig-largest-vancouver-parks
#| fig-cap: "The ten largest public parks in Vancouver by area."

top_ten_parks <- head(largest_parks, 10)

ggplot(
  top_ten_parks,
  aes(
    x = reorder(Name, Hectare),
    y = Hectare
  )
) +
  geom_col(fill = "#457b9d") +
  coord_flip() +
  labs(
    x = "Park",
    y = "Area (hectares)"
  ) +
  theme_minimal()
```

## Findings

The 218 parks in the dataset have a mean area of 6.2 hectares but a median area of only 1.4 hectares. This large difference indicates that the distribution of park sizes is strongly right-skewed: most parks are relatively small, while a few very large parks increase the mean. Stanley Park is the largest at 400 hectares, considerably larger than Fraserview Golf Course, the second-largest site at 84.34 hectares.